---
title: "Using the `hurricaneexposure` package"
author: "Brooke Anderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE}
library(hurricaneexposure)
```

## Loading the package

The package currently exists in [a development version](https://github.com/geanders/hurricaneexposure) on GitHub. You can use the following code to load it:

```{r eval = FALSE}
library(devtools)
install_github("geanders/hurricaneexposure", 
               build_vignettes = TRUE)
library(hurricaneexposure)
```

Once you've loaded the dataframe, you can load the included data using the `data` function. For example:

```{r}
data("hurr_tracks")
head(hurr_tracks)
```

The following datasets are included with the package: 

- `closest_dist`
- `county_centers`
- `hurr_tracks`
- `precip_file`

For each, you can see the helpfiles for the data for more information about the data included in each. 

## Creating time series datasets of exposure

There are several functions in this package that help to take the data on rainfall and storm tracks stored in the package and use them, in conjunction with thresholds for the distance to the storm tracks and the amount of rainfall required for exposure, to create time series of exposure that can be easily integrated into time series datasets of, for example, health data. 

First, the `rain_storms` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds for distance (storm path must come that close or closer to the county) and rainfall (total rainfall for the week window centered on the date that storm was closest to the county must be equal to or above the threshold). The function outputs a dataframe with all of the storms to which each of the counties was exposed.

For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 [units] of the county center and the weekly rainfall was 100 [units] or higher, you could run:

```{r}
county_rain(counties = c("22071", "51700"),
            start_year = 1995, end_year = 2005,
            rain_limit = 100, dist_limit = 100,
            days_included = c(-1, 0, 1))
```

In addition to giving you the names and closest dates of each storm for each county (`closest_date`), this function also gives you the distance between the county and the storm's track at time when the storm was closest to the county center (`storm_dist`) and the total precipitation for the week-long window centered on that closest date (`tot_precip`).

To get "exposure" storms for multi-county communities, you can use the `multi_county_rain` function:

```{r warning = FALSE}
communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
                         fips = c("36005", "36047", "36061",
                                  "36085", "36081", "36119",
                                  "22071", "51700"))
rain_storm_df <- multi_county_rain(communities = communities,
                                   start_year = 1995, end_year = 2005,
                                   rain_limit = 100, dist_limit = 100,
                                   days_included = c(-1, 0, 1))
```

**Note:** You currently will get the following warning when you run this function: 

```
Warning message:
In left_join_impl(x, y, by$x, by$y) :
  joining factor and character vector, coercing into character vector
```

That's not a problem, and I'll change the code later so you don't get this warning.

This output includes columns for the average closest distance for any of the counties in the communities (`mean_dist`), the average of the weekly precipitation for all the counties (`mean_precip`), the highest weekly rain for any of the counties (`max_rain`), and the smallest distance between the storm track and any of the county centers (`min_dist`).

## Creating and writing time series of exposure

To create time series of hurricane exposure for all of the storms that meet the exposure definition for study communities, you can use the function `rain_exposure`. If you have counties for your study, you can run this with a vector of the county FIPS as the `locations` argument (specify the directory path to write the files with `out_dir`):

```{r}
rain_exposure(locations = c("22071", "51700"),
              start_year = 1995, end_year = 2005,
              rain_limit = 100, dist_limit = 100,
              out_dir = "~/tmp/storms")
```

If you have multi-county communities, set `locations` instead to be a dataframe with community names (`commun` column) and FIPS codes (`fips` column):

```{r message = FALSE, warning = FALSE}
communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
                          fips = c("36005", "36047", "36061",
                          "36085", "36081", "36119",
                          "22071", "51700"))
rain_exposure(locations = communities,
              start_year = 1995, end_year = 2005,
              rain_limit = 100, dist_limit = 100, out_dir = "~/tmp/storms")
```

**Note:** You'll get the same message here as you do when running `multi_county_rain`, because this function is running that one. Again, it's not a problem.

## Mapping hurricane exposure

This package allows you to create some different maps of hurricane exposures. 

### Plotting county-level exposure 

The `map_counties` function creates county choropleths of different storm exposure variables (right now, rainfall and distance from the storm tracks). For example, to plot rain exposure for Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):

```{r fig.width = 7, fig.height = 4}
map_1 <- map_counties(storm = "Floyd-1999", metric = "rainfall")
map_1
```

You may notice that this function gives a warning listing the FIPS of any counties for which this exposure information was missing.

You can also use this function to plot the closest distance between the storm and each county. For this, you use the argument `metric = "closest distance"`.

```{r fig.width = 7, fig.height = 4}
map_2 <- map_counties(storm = "Floyd-1999", metric = "distance")
map_2
```

You can map a binary variable of distance-based exposure using `map_distance_exposure`: 

```{r fig.width = 7, fig.height = 4}
allison_map <- map_distance_exposure(storm = "Allison-2001",
                                     dist_limit = 75)
plot(allison_map)
```


You can also map a binary variable of rain exposure for the communities that were exposed, based on a certain rainfall limit and distance limit:

```{r fig.width = 7, fig.height = 4}
map_3 <- map_rain_exposure(storm = "Floyd-1999", rain_limit = 125,
                           dist_limit = 500, 
                           days_included = c(-1, 0, 1))
plot(map_3)
```


### Plotting storm tracks 

The `map_tracks` function will map the hurricane tracks for one or more storms. For example, to plot the tracks of Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):

```{r fig.width = 6.5, fig.height = 4}
map_4 <- map_tracks(storms = "Floyd-1999")
map_4
```

There are some different options you can use for the tracks' appearance. For example, if you wanted to plot the tracks of several storms, not show storm status with color, not plot each point when the track locations were measured (typically every six hours), and use some transparency so you can see all the lines, you can use: 

```{r fig.width = 5.5, fig.height = 4}
map_5 <- map_tracks(storms = c("Floyd-1999", "Sandy-2012",
                               "Katrina-2005"),
                    storm_status = FALSE,
                    plot_points = FALSE,
                    alpha = 0.5)
map_5
```

You can also add these tracks to an existing `ggplot`-created US map. You do this through the `plot_object` argument. For example, to add the storm track to the plot of rain exposure for Floyd, you could run:

```{r fig.width = 7, fig.height = 4}
map_6 <- map_tracks(storms = "Floyd-1999", plot_object = map_1,
                    storm_status = FALSE, plot_points = FALSE)
map_6
```

```{r fig.width = 7, fig.height = 4}
map_7 <- map_tracks(storms = "Floyd-1999", plot_object = map_3,
                    storm_status = FALSE, plot_points = FALSE)
map_7
```


