---
output:
  beamer_presentation:
    keep_tex : true
    includes: 
      in_header: anderson_header.txt
      before_body: anderson_beforebody.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motivation

\vspace{-0.5cm}
\begin{center}
\includegraphics[width = \textwidth]{images/rainexamples.pdf}
\end{center} 
\vspace{-0.8cm}
\textbf{Goal:} Create a publicly-accessible package with historical data (1988--present) on county-level exposure to Atlantic basin tropical storms in the United States (with exposure based on distance to storm track, rain, wind, floods, or tornadoes).

## Motivation

\begin{columns}

\begin{column}{0.45\textwidth}
\begin{block}{``Data" package}
\includegraphics[width = \textwidth]{images/hurrtracks.png} \\
\texttt{hurricaneexposuredata}\\
Hosted with \texttt{drat}. 
\end{block}
\end{column}

\begin{column}{0.45\textwidth}
\begin{block}{``Code" package}
\includegraphics[width = \textwidth]{images/example_hurricane_map.png} \\
\texttt{hurricaneexposure}\\
Available through CRAN. 
\end{block}
\end{column}

\end{columns}

\pause
\bigskip

\begin{block}{Challenges}
\begin{enumerate}
\item \alert<3>{Host a package with \texttt{drat}.} 
\item Connect a CRAN package with a data package hosted with \texttt{drat}.
\end{enumerate}
\end{block}

## Hosting a package with `drat`

\begin{columns}

\begin{column}{0.49\textwidth}

\begin{block}{R repositories and \texttt{drat}}
An R package repository is a network-accessible directory structure, and \texttt{drat} helps you build and maintain your own.
\end{block}

\bigskip

\small

\begin{block}{Extensive resources}
\vspace{-0.3cm}
\begin{itemize}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageUsers.html}{Drat Basics for Package Users}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageAuthors.html}{Drat Basics for Package Authors}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratFAQ.html}{Drat FAQs}
\end{itemize}
\end{block}

\end{column}

\begin{column}{0.49\textwidth}
\includegraphics[width = \textwidth]{images/drat_repo_structure.pdf}
\end{column}

\end{columns}

## Packages hosted on `drat` repositories

### For package users, how is a `drat` package different?

- "Not at all", really.
- Tell R about the additional repository via
    - repos argument directly in `install.packages()` et al.,
    - setting another `repos` field in `options()`, or
    - using `drat` helper functions `addRepo()` or `drat:::add()`
- Full details about `repos` in `help(download.packages)`

### For package maintainers, how is a `drat` package different?

- Need to create a personal repository and make it available to users
- Can use functions in the `drat` package to manage this repository 

## Coding a CRAN package to use a `drat` package

### How is GitHub involved?

- Hosting on GitHub particularly easy because 
    - Every GitHub repo has an (optional) web presence
    - A GitHub username plus `drat` is a unique URL
- So if we assume `drat` as the (GitHub) repository name, then we can reference it with just the username:  `addRepo("geanders")` 

## Coding a CRAN package to use a `drat` package

\begin{columns}

\begin{column}{0.45\textwidth}
\begin{block}{``Data" package}
\includegraphics[width = \textwidth]{images/hurrtracks.png} \\
\textit{`hurricaneexposuredata`}\\
Host with \texttt{drat}. 
\end{block}
\end{column}

\begin{column}{0.45\textwidth}
\begin{block}{``Code" package}
\includegraphics[width = \textwidth]{images/example_hurricane_map.png} \\
\textit{`hurricaneexposure`}\\
Available through CRAN. 
\end{block}
\end{column}

\end{columns}

\bigskip

\begin{block}{Challenges}
\begin{enumerate}
\item Host a package with \texttt{drat}. 
\item \alert{Connect a CRAN package with a data package hosted with \texttt{drat}.}
\end{enumerate}
\end{block}

## Technical support experience, non-R

\begin{quotation}
``Unfortunately it is out of our area of expertise to do deep level troubleshooting on an individual's computer to find the cause of the issue. In this case I would recommend using a colleague's computer and see if that is successful in creating a submission."\\[0.5ex]
-- Example of recent non-R technical support 
\end{quotation}

## Technical support experience, R

\begin{center}
\begin{enumerate}
\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/github_search.png}}
          \end{minipage} \medskip \pause
          
\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/the_initial_tweet.png}}
          \end{minipage} \medskip \pause
          
\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/r_journal_title.png}}
          \end{minipage}
          
\end{enumerate}
\end{center}

## Coding a CRAN package to use a `drat` package

### How hard is it to code a CRAN package to "depend" on a `drat` package?

- Not very
- The `DESCRIPTION` file has an `Additional_repositories` field
     - Optional packages can reside on `Additional_repositories`
     - Mandatory packages (`Imports:`, `Depends:`, `LinkingTo:`) cannot
     - But `Suggests:` can point there -- feature we use
- Repos created with `drat` are a good use case for `Additional_repositories`
- \alert{But} also ensure you test for presence of optional package

## Coding a CRAN package to use a `drat` package

### How hard is it to code a CRAN package to "depend" on a `drat` package?

- Need to add code in the CRAN package to do three jobs: 
    + Add infrastructure in the DESCRIPTION file
    + Customize behavior when the package is loaded 
    + Condition code based on availability of `drat` package (functions, vignette, examples)
  
## Coding a CRAN package to use a `drat` data package

\begin{block}{Add infrastructure in the DESCRIPTION file}
\begin{itemize}
\item Include the name of the data package in the ``Suggests" field.
\item Give the address of the \texttt{drat} repository with the data package in the ``Additional\_repositories" field.
\end{itemize}
\end{block}

\bigskip

For example, the `hurricaneexposure` DESCRIPTION file includes: 
\vspace{-1cm}

```
Suggests: hurricaneexposuredata (>= 0.0.2)
Additional_repositories: https://geanders.github.io/drat
```

## Coding a CRAN package to use a `drat` data package

\begin{block}{Customize behavior when the package is loaded }
\begin{itemize}
\item These steps can be included in ``zzz.R" (or similar) file, by writing \texttt{.onLoad} and \texttt{.onAttach} functions.
\item Check if the user has the data package and create an environment that stores the result. 
\item If the user does not have the data package, give a start-up message telling the user how to install it. 
\end{itemize}
\end{block}

## Coding a CRAN package to use a `drat` data package

\begin{block}{Condition code based on availability of data package}
\begin{itemize}
\item Condition code in functions that need access to the data package (e.g., helper function can be written in the package).
\item Condition code in documentation examples to only run if the data package is available.
\item Condition vignette to only evaluate code chunks if the data package is available.
\end{itemize}
\end{block}

## Coding a CRAN package to use a `drat` data package

### Example: Conditioning code in the vignette

A code chunk like the following can be included at the start of the vignette to condition the evaluation of code chunks on having the data package available:

\footnotesize

```{r eval = FALSE, echo = TRUE}
hasData <- requireNamespace("hurricaneexposuredata", quietly = TRUE)
if (!hasData){
  knitr::opts_chunk$set(eval = FALSE)
  msg <- paste("Note: Examples in this vignette require that the",
               "`hurricaneexposuredata` package be installed. The",
               "system currently running this vignette does not",
               "have that package installed, so code examples",
               "will not be evaluated.")
  msg <- paste(strwrap(msg), collapse = "\n")
  message(msg)
}
```


## Find out more

\footnotesize

\begin{block}{Article in The R Journal}
 \href{https://journal.r-project.org/archive/2017/RJ-2017-026/index.html}{Anderson and Eddelbuettel. 2017. Hosting Data Packages via drat: A Case Study with Hurricane Exposure Data. \textit{The R Journal}. 9(1):486--497.}
\end{block}

\begin{block}{Example R packages}
\begin{itemize}
\item \texttt{hurricaneexposure}: \href{www.github.com/geanders/hurricaneexposure}{geanders/hurricaneexposure}
\item \texttt{hurricaneexposuredata}: \href{www.github.com/geanders/hurricaneexposuredata}{geanders/hurricaneexposuredata}
\end{itemize}
\end{block}

\begin{block}{\texttt{drat} documentation}
\begin{itemize}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageUsers.html}{Drat Basics for Package Users}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageAuthors.html}{Drat Basics for Package Authors}
\end{itemize}
\end{block}

\begin{block}{Acknowledgments}
Thanks to Martin Morgan and Roger Peng for helpful suggestions. This work was supported in part by a grant from the National Institute of Environmental Health Sciences (R00ES022631)
\end{block}
